Simulation Learning in Computer Hex

### 2011-12-06 thesis/thesis.tex
Added appendix A1

200 games per pair were played, with time controls set at 5 seconds per move. All players have approximately the same performance with regard to simulations per second, so the 5 seconds per move resulted in around 10,000 simulated playouts. A boardsize of 7x7 was used, and to control for the swap rule a swap-safe move was used as the first move for black. Matches were run using the Gomill software\cite{gomill}. Gomill required no modification to work with Hex playing programs. Results are shown in figure \ref{fig:results}.

@@

>>>>

200 games per pair were played, with time controls set at 10,000 playouts per move. A boardsize of 7x7 was used, and to control for the swap rule a swap-safe move was used as the first move for black. Matches were run using the Gomill software\cite{gomill}. Gomill required no modification to work with Hex playing programs. Results are shown in figure \ref{fig:results}.

@@

\section{Appendix}\label{appendix}
\subsection{A1 - Stability of Evolution}
Stochastic processes like evolutionary search can have large variations in their end results. To test the variation of Hivemind, 10 independent individuals were produced. Each individual was the best individual from the 100th generation of an independent run of the learning process. The parameters used for each run were that same as those used for the final results of self-play,  \texttt{generations=100, lambda=40, mu=30, parents=2, samples=5}, with 1000 playouts per move on a 7x7 board. To test, each independent individual (Learned 1 through 10) played 200 games on a 7x7 board, against a player using the default policy (Default), all players simulating 10,000 playouts per move.

\begin{figure}
	\begin{center}
		\begin{tabular}{c c c c}
		& Default \\
		\hline
		Learned 1 & 91.0\% \\
		Learned 2 & 90.5\% \\
		Learned 3 & 90.5\% \\
		Learned 4 & 87.5\% \\
		Learned 5 & 90.5\% \\
		Learned 6 & 93.5\% \\
		Learned 7 & 93.0\% \\
		Learned 8 & 91.5\% \\
		Learned 9 & 87.0\% \\
		Learned 10 & 92.5\% \\
		Min & 87.0\% \\
		Max & 93.5\% \\
		Mean & 90.75\% \\
		Std. Dev. & ~2.14\% \\
		\end{tabular}
		\caption{Results of 10 independent evolutionary processes versus the default policy.}
		\label{fig:a1results}
	\end{center}
\end{figure}

Figure \ref{fig:a1results} shows the results of these games. Overall, the 10 individuals had a standard deviation of about 2.14\%, with a mean of 90.75\%. This indicated that the evolutionary processes is relatively stable. With only a few independent runs of the evolutionary processes, an implementer can reasonably assume that the best individual would perform as well as after several hundred runs.

\subsection{A2 - Effects of different board sizes}
The computational time required of learning process greatly depends on the boardsize used in learning. For training, a boardsize of 7x7 was chosen as a tradeoff between computation time and bias toward first-mover advantage.
